### 2018-04-29 Paper/bg.tex
..

The derivative-based global sensitivity analysis (DGSA) is performed by 
estimating a sensitivity measure~\cite{Kucherenko:2009} for each stochastic
parameter in a model or a function. The
sensitivity measure associated with a parameter, $\theta_i$, based on
the observable, $\mathcal{Y}$ is given as follows:

@@

The derivative in the above equation is estimated using finite
difference as given below, and averaged using pseudo-random samples in
the uncertain parameter domain.

@@

The upper bound in the above inequality is proportional to the product of $\mathcal{C}_i$
and $\mu_i$. However, for the purpose of parameter screening as discussed later in
Section~\ref{sec:method}, we consider a normalized product, $\hat{\mathcal{C}_i\mu_i}$:

@@

The Poincar\'e constant, $\mathcal{C}_i$ is specific to the probability distribution of $\theta_i$.
Table~\ref{tab:poincare} provides its value in the case of uniform and normal
distributions.
\bigskip

>>>>

The derivative-based global sensitivity analysis (DGSA) is performed by 
estimating a sensitivity measure~\cite{Sobol:2009} for each stochastic
parameter in a model or a function. The
sensitivity measure associated with a parameter, $\theta_i$, based on
the observable, $\mathcal{Y}$ is given as follows:

@@

\noindent The derivative in the above equation is estimated using finite
difference as given below, and averaged using pseudo-random samples in
the uncertain parameter domain.

@@

\noindent The upper bound in the above inequality is proportional to the product of $\mathcal{C}_i$
and $\mu_i$. However, for the purpose of parameter screening as discussed later in
Section~\ref{sec:method}, we consider a normalized product, $\hat{\mathcal{C}_i\mu_i}$:

@@

\noindent The Poincar\'e constant, $\mathcal{C}_i$ is specific to the probability distribution of $\theta_i$.
Table~\ref{tab:poincare} provides its value in the case of uniform and normal
distributions.
\bigskip

### 2018-04-29 Paper/examples.tex
update ...
..

As mentioned earlier in this section, cheap function evaluations of the discharge using 
Eq.~\ref{eq:bore} enable us to compute converged estimates of $\mathcal{T}(\theta_i)$. Plotted in
Figure~\ref{fig:sense_bore} are estimates of these indices corresponding to the uncertain
parameters in the borehole function using 10$^6$ pseudo-random
samples\footnote{Although $\mathcal{T}(\theta_i)$ 
might converge with much fewer samples depending upon the model, we consider a large number that
typically ensures a converged estimate for the purpose of illustration.} in the input parameter domain. 
These estimates are used to verify fidelity of
parameter screening based on the methodology presented in Section~\ref{sec:method}. 
In Figure~\ref{fig:screen_bore}, we plot estimates of the screening parameter ($\hat{\mathcal{C}_i\mu_i}$)
for a wide range of the number of samples used for approximating $\mu_i$ using Eq.~\ref{eq:mu}.
Estimates for $\hat{\mathcal{C}_i\mu_i}$ are found to be in excellent agreement with $\mathcal{T}(\theta_i)$
even when small number of samples (5 -- 10) are used. 
Consequently, the relative importance of uncertain 
parameters in the borehole function is found to be consistent with predictions based on the Sobol' index. 
In the considered intervals for the uncertain parameters, it is clear that the discharge is insensitive to
$T_u$ and $T_l$. Moreover, the sensitivity towards $K_w$ is also small. We exploit this result to reduce
the dimensionality of the problem by 
discounting the uncertainties in $T_u$, $T_l$, and $K_w$, while using nominal values for these
parameters instead.

Our goal as discussed is to gain computational advantage by constructing surrogates in a 
reduced input parameter space. To this end, we construct PCE's in 5D and 4D spaces by
fixing $\{T_u,T_l\}$ in the former and additionally fixing $K_w$ in the latter at their respective
mean values. In Figure~\ref{fig:conv_bore}, we compare convergence of PCE's constructed
in the full space (7D) with those constructed in the two reduced spaces (4D and 5D) using
$\epsilon_{\mbox{\tiny{LOO}}}$ (Eq.~\ref{eq:loo}). As expected, it is observed that the PCE
constructed in the 4D space converges at a much faster rate. For instance, if a PCE with
$\mathcal{O}(10^{-4})$ accuracy is sought, we need function evaluations at only about
50 sample points in the 4D parameter space whereas the number of samples needed in the full 
7D space seems much higher.

@@

Figure~\ref{fig:pdfcomp_bore} illustrates a comparison of the PDF's of the discharge, $\mathcal{Q}$
obtained by propagating 10$^6$ Monte-Carlo (MC) samples through the 7D PCE in the
original input parameter domain as well as the reduced order PCE's constructed in 4 and 5 dimensions. 
It is evident from this plot, that the PDF's agree quite favorably with each other. Consequently, it can
be said that the reduced order PCE is verified in a probabilistic sense. In other words, the mode as
well as the uncertainty in the observable is reliably captured and predicted by the reduced order PCE.

@@

>>>>

As mentioned earlier in this section, cheap function evaluations of the discharge using 
Eq.~\ref{eq:bore} enable us to compute converged estimates of $\mathcal{T}(\theta_i)$. Plotted in
Figure~\ref{fig:sense_bore} are estimates of these indices corresponding to the uncertain
parameters in the borehole function using 10$^6$ pseudo-random
samples\footnote{Although $\mathcal{T}(\theta_i)$ 
might converge with much fewer samples depending upon the model, we consider a large number that
typically ensures a converged estimate for the purpose of illustration.} in the input parameter domain. 
These estimates are used to verify fidelity of
parameter screening based on the methodology presented in Section~\ref{sec:method}.

\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.65\textwidth]{./Figures/sense_borehole}
\caption{Sobol' total effect sensitivity index for uncertain parameters in the borehole
function (Eq.~\ref{eq:bore}).}
\label{fig:sense_bore}
\end{center}
\end{figure}

In Figure~\ref{fig:screen_bore}, we plot estimates of the screening parameter ($\hat{\mathcal{C}_i\mu_i}$)
for a wide range of the number of samples used for approximating $\mu_i$ using Eq.~\ref{eq:mu}.

\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.65\textwidth]{./Figures/ub_conv_borehole}
\caption{Estimates of the screening parameter ($\hat{\mathcal{C}_i\mu_i}$) is plotted 
against number of samples considered for evaluating $\mu_i$. Corresponding estimates
for the converged Sobol' total effect index ($\mathcal{T}(\theta_i)$) in each case is also included in the
legend.}
\label{fig:screen_bore}
\end{center}
\end{figure}

\noindent Estimates for $\hat{\mathcal{C}_i\mu_i}$ are found to be in excellent agreement with $\mathcal{T}(\theta_i)$
even when small number of samples (5 -- 10) are used. 
Consequently, the relative importance of uncertain 
parameters in the borehole function is found to be consistent with predictions based on the Sobol' index. 
In the considered intervals for the uncertain parameters, it is clear that the discharge is insensitive to
$T_u$ and $T_l$. Moreover, the sensitivity towards $K_w$ is also small. We exploit this result to reduce
the dimensionality of the problem by 
discounting the uncertainties in $T_u$, $T_l$, and $K_w$, while using nominal values for these
parameters instead.

Our goal as discussed is to gain computational advantage by constructing surrogates in a 
reduced input parameter space. To this end, we construct PCE's in 5D and 4D spaces by
fixing $\{T_u,T_l\}$ in the former and additionally fixing $K_w$ in the latter at their respective
mean values. In Figure~\ref{fig:conv_bore}, we compare convergence of PCE's constructed
in the full space (7D) with those constructed in the two reduced spaces (4D and 5D) using
$\epsilon_{\mbox{\tiny{LOO}}}$ (Eq.~\ref{eq:loo}).

\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.65\textwidth]{./Figures/err_samples_borehole}
\caption{A comparison of order of the leave-one-out-error 
($\epsilon_{\mbox{\tiny{LOO}}}$) as a function of number of regression samples
used for constructing the PCE in 4, 5, and 7 dimensions.}
\label{fig:conv_bore}
\end{center}
\end{figure}

\noindent As expected, it is observed that the PCE
constructed in the 4D space converges at a much faster rate. For instance, if a PCE with
$\mathcal{O}(10^{-4})$ accuracy is sought, we need function evaluations at only about
50 sample points in the 4D parameter space whereas the number of samples needed in the full 
7D space seems much higher.

@@

Figure~\ref{fig:pdfcomp_bore} illustrates a comparison of the PDF's of the discharge, $\mathcal{Q}$
obtained by propagating 10$^6$ Monte-Carlo (MC) samples through the 7D PCE in the
original input parameter domain as well as the reduced order PCE's constructed in 4 and 5 dimensions.

\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.65\textwidth]{./Figures/pdf_comp_borehole}
\caption{A comparison of PDF's of the discharge, $\mathcal{Q}$ as obtained
by propagating
parametric uncertainties through the 4D, 5D, and 7D PCE's.These PDF's 
were obtained using 10$^{6}$ MC samples in the input parameter domain.} 
\label{fig:pdfcomp_bore}
\end{center}
\end{figure}

\noindent It is evident from this plot, that the PDF's agree quite favorably with each other. Consequently, it can
be said that the reduced order PCE is verified in a probabilistic sense. In other words, the mode as
well as the uncertainty in the observable is reliably captured and predicted by the reduced order PCE.

@@

\subsection{Semilinear elliptic PDE with random source term}

We consider the semilinear elliptic PDE,
\begin{equation}\label{equ:semilinear}
\begin{aligned}
-\kappa \Delta u + c u^3 &= q, \quad \text{in } \Omega\\
 u &= 0. \quad \text{on } \partial \Omega
\end{aligned}
\end{equation}
We model the source term as 
\begin{equation}
q(x, y, \theta) = \sum_{i = 1}^N \theta_i \sin(i \pi x)
\cos(i \pi y),
\quad N = 8.
\end{equation}
Here $\kappa \sim U(0.08, 0.12)$, $c \sim U(0.8, 1.2)$, 
and $\theta_i$ are uniformly distributed in interval $[0, 1]$. We consider
the following QoI,
\[
     f(\kappa, c, \theta) = \frac{1}{|D|} \int_D u(x; \kappa, c, \theta) \, dx, 
\]
where $D$ is the region, $[1/2, 3/4] \times [1/2, 3/4]$.

### 2018-04-29 Paper/intro.tex
..

The emerging field of uncertainty quantification (UQ) aims at methodologies for 
incorporating, characterizing, quantifying, propagating, and reducing the 
uncertainties associated with predictive models and simulations. For situations
involving complex physical models and compute-intensive simulations, an
efficient approach to construction of model surrogates is sought to enable UQ
in a tractable manner. Polynomial chaos expansion 
(PCE)~\cite{Xiu:2002,Ghanem:2003,Eldred:2008,Olivier:2010} and 
Gaussian Process (GP)~\cite{Rasmussen:2004} or Kriging~\cite{Stein:2012} are
among the most commonly used surrogates for scientific applications. However,
both approaches quickly become prohibitive when the set of uncertain model
inputs and parameters is large-dimensional. Specifically, in the case of a PCE,
a $d$-dimensional polynomial basis with a total order of truncation, $p$ 
requires model realizations at $(p+1)^d$ quadrature nodes to estimate the PC
coefficients using a fully-tensorized Gauss quadrature. This trend is 
illustrated below in Figure~\ref{fig:curse}.

\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.70\textwidth]{./Figures/quad_comp}
\caption{Number of fully-tensorized Gauss quadrature nodes ($(p+1)^d$) is plotted against PCE total order truncation, $p$ for $d$ = 2,3, and 4.}
\label{fig:curse}
\end{center}
\end{figure}

>>>>

The emerging field of uncertainty quantification (UQ) aims at methodologies for 
incorporating, characterizing, quantifying, propagating, and reducing the 
uncertainties associated with predictive models and simulations. For situations
involving complex physical models and compute-intensive simulations, an
efficient approach to construction of model surrogates is sought to enable UQ
in a tractable manner. Polynomial chaos expansion 
(PCE)~\cite{Xiu:2002,Ghanem:2003,Eldred:2008,Olivier:2010} and 
Gaussian Process (GP)~\cite{Rasmussen:2004} or Kriging~\cite{Stein:2012} are
among the most commonly used surrogates for scientific applications. However,
both approaches quickly become prohibitive when the set of uncertain model
inputs and parameters is large-dimensional. Specifically, in the case of a PCE,
a $d$-dimensional polynomial basis with a total order of truncation, $p$ 
requires model realizations at $(p+1)^d$ quadrature nodes to exactly determine the PC
coefficients using a fully-tensorized Gauss quadrature. Consider a case where $d$ = 8 which
could be a regarded as a low-dimensional problem, and $p$ = 5, requires model evaluations at
roughly 1.7 million quadrature nodes. However, the degree of exactness in numerical estimation 
of the coefficients as made possible with a fully-tensored quadrature might not be desired. In that
case, computational effort can be reduced considerably by using nested
quadratures~\cite{Gentleman:1972a,Gentleman:1972b,Novak:1999,Waldvogel:2006} which allow 
model realizations at a given level to be used at higher levels of accuracy as well as sparse grids
based on Smolyak tensorization~\cite{Smolyak:1963}. Additonally, sparse basis 
techniques~\cite{Peng:2014,Hampton:2015,Blatman:2011} have been used to reduce  the required
computational effort for constructing a reasonably accurate PC surrogate.

One of the primary objectives for constructing a model surrogate is to be able to perform 
parametric sensitivity analysis tractably. Sensitivity analysis provides an insight into the
relative contributions of the uncertain model parameters and inputs to the uncertainty in
predictions. The analysis could potentially reduce the dimensionality of the problem.
However, as discussed earlier, constructing the surrogate might still require a large amount
of computational resources. It can thus be understood that dimension reduction prior to
constructing the surrogate might lead to enormous computational savings. In this work,
we present a strategy for identifying and screening uncertain model parameters that 
are significantly less important that the rest, thereby reducing the dimensionality of the
problem and enabling the construction of a reduced-order surrogate. As discussed in
section~\ref{sec:method}, parameter ranks are determined using a \textit{screening-metric}
based on derivative-based global sensitivity 
measures~\cite{Sobol:2009,Sobol:2010,Lamboni:2013,Kucherenko:2009,Kucherenko:2016}. 
The strategy is agnostic to the choice of methodology for constructing the surrogate. However, 
we will be using PCE to demonstrate the suitability of the proposed strategy. Below, in 
section~\ref{sec:bg}, we provide a brief introduction to derivative-based global sensitivity analysis
in the present context as well as the polynomial chaos methodology. Later, in section~\ref{sec:examples},
we motivate the proposed strategy by implementing it to a suite of simple applications, namely, the 
classic borehole function, a non-linear oscillator, and a semi-linear elliptic PDE. Finally, we demonstrate
suitability of our approach by using it to reduce the dimensionality of a H$_2$/O$_2$ kinetics problem
involving 19 reactions in section~\ref{sec:app}.

%\begin{figure}[htbp]
% \begin{center}
%  \includegraphics[width=0.70\textwidth]{./Figures/curse}
%\caption{Number of multi-dimensional Gauss quadrature nodes, $(p+1)^d$ is plotted against the PCE
% total order truncation, $p$ for 2, 3, and 4 dimensional polynomial basis functions.}
%\label{fig:curse}
%\end{center}
%\end{figure}
